---
title: "OSRI.Figures"
output: html_document
date: "2025-08-28"
editor_options: 
  chunk_output_type: console
---
Set Up
```{r setup, include=FALSE}

library(viridis)
library(ggstream)

D19<-read_xlsx("OSRI_Totals_Full.xlsx")

D19<-D19%>%#Reorder layman group to be more intuitive
  mutate(Layman_group = factor(Layman_group, levels = c("Seaweed","Snails","Urchins/ Starfish","Octopus","Clams", "Mussels","Crabs/ Shrimp","Fish","Seals/ Whales")))
```


Figure 1 - Data in Analysis- Stacked Area Chart with Study Observations over Time
```{r setup, include=FALSE}
all_years <- unique(D19$Year_clean)
all_studies <- unique(D19$Study_name_cleaner)
complete_grid <- expand.grid(Year_clean = all_years, Study_name_cleaner = all_studies)

summary_df <- D19 %>%
  group_by(Year_clean, Study_name_cleaner) %>%
  summarise(Count = n(), .groups = "drop")

summary_df_complete <- complete_grid %>%
  left_join(summary_df, by = c("Year_clean", "Study_name_cleaner")) %>%
  mutate(Count = replace_na(Count, 0)) %>%
  arrange(Year_clean, Study_name_cleaner)

summary_cumulative <- summary_df_complete %>%
  group_by(Study_name_cleaner) %>%
  arrange(Year_clean) %>%
  mutate(CumulativeCount = cumsum(Count)) %>%
  ungroup()%>%
  mutate(
    Year_clean=as.numeric(Year_clean)
  )

# Plot
ggplot(summary_cumulative, aes(x = Year_clean, y = Count, group = Study_name_cleaner, fill= Study_name_cleaner)) +
  geom_stream(color = "white", size = 0.1, alpha = 0.9,type = "ridge", bw=1) +
  scale_fill_viridis_d(name = "Study",option = "D") +
  theme_minimal(base_size = 13) +
  labs(
    title = "",
    x= "",
    y = "Number of Observations") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = c(1.025, 1.10),  # Position legend inside plot (x, y from bottom-left)
    legend.justification = c("right", "top"),  # Anchor the legend box
    legend.background = element_blank(),   # transparent background
    legend.box.background = element_blank(),  # remove outer box
    legend.key = element_blank(),           # transparent boxes behind keys
    legend.title = element_text(face = "bold"))

ggsave(filename = "Figures/Fig1-ObservationsByStudyOverTime.png",    # output file name
  plot = last_plot(),                   # or specify your plot object like `my_plot`
  device = "png",
  width = 8.5,                          # half of standard 11-inch paper width
  height = 5.5,                         # full height of standard paper
  units = "in",
  dpi=300)
```

##Figure 2a - Pie chart of taxa and observations with list of species under each label

```{r setup, include=FALSE}
library(ggrepel)

taxa_summary <- D19 %>%
  group_by(Layman_group, Common_name) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(Layman_group) %>%
  summarise(
    TotalCount = sum(Count),
    SpeciesList = paste(unique(Common_name), collapse = ", "),
    .groups = "drop"
  ) %>%
  arrange(desc(TotalCount)) %>%
  mutate(
    Fraction = TotalCount / sum(TotalCount),
    ymax = cumsum(Fraction),
    ymin = lag(ymax, default = 0),
    label_pos = (ymax + ymin)/5,
    Label = paste0(Layman_group, "\n(n=", TotalCount, ")")
  )
write_xlsx(taxa_summary, "Tables/taxa_summary_Figure2.xlsx")
# Step 2: Create Plot
ggplot(taxa_summary, aes(x = "", y = TotalCount, fill = Layman_group)) +
  geom_bar(stat = "identity", position = "stack",width = 10) +
  coord_polar(theta = "y", direction = 1) +
  theme_void() +
  scale_fill_viridis_d(name = "Taxonomic Group") 
#Silly Labels that never worked
#####
  # Internal pie labels
  geom_text(
    aes(x = label_pos, y = label_pos, label = Label),
    color = "white",
    size = 3,
    fontface = "bold"
  ) #+

  # External species annotations
  geom_label_repel(
    aes(x = 1.8, y = label_pos, label = SpeciesList, fill = Layman_group),
    size = 2.8,
    box.padding = 0.3,
    segment.size = 0.4,
    segment.color = "gray60",
    label.size = 0.2,
    force = 1.5,
    nudge_x = 0.5,
    direction = "y",
    max.overlaps = 200,
    show.legend = FALSE
  ) +

  # Make sure labels fit
  #xlim(0.5, 2.5) +
  theme(legend.position = "none")
#####
ggsave(filename = "Figures/Fig2a-SpeciesObservationsBreakdown.png",    # output file name
  plot = last_plot(),                   # or specify your plot object like `my_plot`
  device = "png",
  width = 8.5,                          # half of standard 11-inch paper width
  height = 5.5,                         # full height of standard paper
  units = "in",
  dpi=300)
#####

#2b- Stacked Barchart by Observations by Region
  
#####
missing_rows <- D19 %>%
  filter(is.na(Lat)) # Yes, some stations are missing location Data
  
D20<-read_csv("OSRI_Totals_region.csv")
D20<-D20%>%#Reorder layman group to be more intuitive
  mutate(Layman_group = factor(Layman_group, levels = c("Seaweed","Snails","Urchins/ Starfish","Octopus","Clams", "Mussels","Crabs/ Shrimp","Fish","Seals/ Whales")))%>%#Reorder region to be more intuitive
  mutate(Region = factor(Region, levels= c("SEAK","SCAK","Aleutian","West","North")))

taxa_summary1 <- D20 %>%
  group_by(Layman_group, Common_name, Region) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(Layman_group, Region) %>%
  summarise(
    TotalCount = sum(Count),
    SpeciesList = paste(unique(Common_name), collapse = ", "),
    .groups = "drop"
  ) %>%
  arrange(desc(TotalCount))
write_xlsx(taxa_summary1, "taxa_summary_Figure2b_Region.xlsx")

ggplot(taxa_summary1, aes(x = Region, y = TotalCount, fill = Layman_group)) +
  geom_bar(stat = "identity", position = "stack") +
   labs(
    title = "",
    x= "Region",
    y = "Number of Observations") +
  theme_minimal() +
  scale_fill_viridis_d(name = "Taxonomic Group") 

ggsave(filename = "Figures/Fig2b-SpeciesObservationsByRegion.png",    # output file name
  plot = last_plot(),                   # or specify your plot object like `my_plot`
  device = "png",
  width = 8.5,                          # half of standard 11-inch paper width
  height = 5.5,                         # full height of standard paper
  units = "in",
  dpi=300)
#####

#Figure 3 - Boxplots of Taxa and all observations Total PAHs

#####
ggplot(data=D19, aes(x = `Layman_group`, y = log(`Total_PAH`), color = `Layman_group`)) +
  geom_boxplot(outlier.shape = NA) +  # Removes the outlier markers from the boxplot
  geom_jitter(aes(color = `Layman_group`), 
              width = 0.2, height = 0, size = 1, alpha = 0.5) +  # Add jittered points
  labs(title = "Log PAH concentrations measured in marine organisms", x = "", y = "log(Total PAHs (ng/g))") +
  theme_minimal() +
  theme(
    legend.position = "none",  # Removes the legend
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis text
  )

ggsave(filename = "Figures/Fig3-LogMeasurementsByTaxa.png",    # output file name
  plot = last_plot(),                   # or specify your plot object like `my_plot`
  device = "png",
  width = 8.5,                          # 
  height = 5.5,                         # 
  units = "in",
  dpi=300)

###Non Log version
ggplot(data=D19, aes(x = `Layman_group`, y = `Total_PAH`, color = `Layman_group`)) +
  geom_boxplot(outlier.shape = NA) +  # Removes the outlier markers from the boxplot
  geom_jitter(aes(color = `Layman_group`), 
              width = 0.2, height = 0, size = 1, alpha = 0.5) +  # Add jittered points
  labs(title = "PAH concentrations measured in marine organisms", x = "", y = "Total PAHs (ng/g)") +
  theme_minimal() +
  theme(
    legend.position = "none",  # Removes the legend
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis text
  )

ggsave(filename = "Figures/Fig3-MeasurementsByTaxa.png",    # output file name
  plot = last_plot(),                   # or specify your plot object like `my_plot`
  device = "png",
  width = 8.5,                          # half of standard 11-inch paper width
  height = 5.5,                         # full height of standard paper
  units = "in",
  dpi=300)

##Clipped Version
ggplot(data=D19, aes(x = `Layman_group`, y = `Total_PAH`, color = `Layman_group`)) +
  geom_boxplot(outlier.shape = NA) +  # Removes the outlier markers from the boxplot
  geom_jitter(aes(color = `Layman_group`), 
              width = 0.2, height = 0, size = 1, alpha = 0.5) +  # Add jittered points
  labs(title = "PAH concentrations measured in marine organisms", x = "", y = "Total PAHs (ng/g)") +
  theme_minimal() +
  ylim(0,250)+
  theme(
    legend.position = "none",  # Removes the legend
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis text
  )

ggsave(filename = "Figures/Fig3-Clipped250MeasurementsByTaxa.png",    # output file name
  plot = last_plot(),                   # or specify your plot object like `my_plot`
  device = "png",
  width = 8.5,                          # half of standard 11-inch paper width
  height = 5.5,                         # full height of standard paper
  units = "in",
  dpi=300)
#####

#Figure 4 - Panel break down of Results in Fish, inverts, and marine mammal by species_standard
library(dplyr)
library(ggplot2)
library(forcats)

##Fish
#####
custom_order <- c("salmon", "anadromous/freshwater fish", "forage fish", "flatfish", "cod", "other fish")

D19_fish <- D19 %>%
  filter(Layman_group == "Fish") %>%
  mutate(
    logPAH = log(Total_PAH),
    Species_standard = factor(Species_standard, levels = custom_order)  # Custom facet order
  ) %>%
  group_by(Species_standard, Common_name) %>%
  mutate(Common_name = fct_reorder(Common_name, logPAH, .fun = median, .desc = TRUE)) %>%
  ungroup()


ggplot(D19_fish, aes(x = Common_name, y = logPAH, color = Common_name)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2, size = 1, alpha = 0.5) +
  facet_wrap(~Species_standard, scales = "free_x"
             , nrow = 1, labeller = label_wrap_gen(width = 15) ) +
  labs(
    title = "Log PAH concentrations measured in Fish",
    x = "",
    y = "log(Total PAHs (ng/g))"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text( size = 9)
  )
ggsave(filename = "Figures/Fig4a-LogMeasurementsinFish.png",    # output file name
  plot = last_plot(),                   # or specify your plot object like `my_plot`
  device = "png",
  width = 10,                          # half of standard 11-inch paper width
  height = 4.5,                         # full height of standard paper
  units = "in",
  dpi=300)

#Non Log Version
D19_fish <- D19 %>%
  filter(Layman_group == "Fish") %>%
  mutate(
    PAH = Total_PAH,
    Species_standard = factor(Species_standard, levels = custom_order)  # Custom facet order
  ) %>%
  group_by(Species_standard, Common_name) %>%
  mutate(Common_name = fct_reorder(Common_name, PAH, .fun = median, .desc = TRUE)) %>%
  ungroup()


ggplot(D19_fish, aes(x = Common_name, y = PAH, color = Common_name)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2, size = 1, alpha = 0.5) +
  facet_wrap(~Species_standard, scales = "free_x"
             , nrow = 1, labeller = label_wrap_gen(width = 15) ) +
  labs(
    title = "PAH concentrations measured in Fish",
    x = "",
    y = "Total PAHs (ng/g)"
  ) +
  theme_minimal() +
  ylim(0, 150) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text( size = 9)
  )
ggsave(filename = "Figures/Fig4a-Clipped150MeasurementsinFish.png",    # output file name
  plot = last_plot(),                   # or specify your plot object like `my_plot`
  device = "png",
  width = 10,                          # half of standard 11-inch paper width
  height = 4.5,                         # full height of standard paper
  units = "in",
  dpi=300)
#####

##Inverts
#####
custom_order <- c("Snails","Urchins/Starfish","Octopus","Clams","Mussels","Crabs/ Shrimp")

D19_inverts <- D19 %>%
  filter(Layman_group %in% c("Mussels","Crabs/ Shrimp","Urchins/Starfish","Octopus","Snails","Clams")) %>%
  mutate(
    logPAH = log(Total_PAH),
    Layman_group = factor(Layman_group, levels = custom_order)  # Custom facet order
  ) %>%
  group_by(Layman_group, Common_name) %>%
  mutate(Common_name = fct_reorder(Common_name, logPAH, .fun = median, .desc = TRUE)) %>%
  ungroup()


ggplot(D19_inverts,  aes(x = Common_name, y = logPAH, color = Common_name)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2, size = 1, alpha = 0.5) +
  facet_wrap(~Layman_group, scales = "free_x"
             , nrow = 1, labeller = label_wrap_gen(width = 15) ) +
  labs(
    title = " Log PAH concentrations measured in Invertebrates",
    x = "",
    y = "log(Total PAHs (ng/g))"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text( size = 9)
  )
ggsave(filename = "Figures/Fig4b-LogMeasurementsinInverts.png",    # output file name
  plot = last_plot(),                   # or specify your plot object like `my_plot`
  device = "png",
  width = 8.5,                          # half of standard 11-inch paper width
  height = 4.5,                         # full height of standard paper
  units = "in",
  dpi=300)

D19_inverts <- D19 %>%
  filter(Layman_group %in% c("Mussels","Crabs/ Shrimp","Urchins/Starfish","Octopus","Snails","Clams")) %>%
  mutate(
    PAH = Total_PAH,
    Layman_group = factor(Layman_group, levels = custom_order)  # Custom facet order
  ) %>%
  group_by(Layman_group, Common_name) %>%
  mutate(Common_name = fct_reorder(Common_name, PAH, .fun = median, .desc = TRUE)) %>%
  ungroup()


ggplot(D19_inverts,  aes(x = Common_name, y = PAH, color = Common_name)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2, size = 1, alpha = 0.5) +
  facet_wrap(~Layman_group, scales = "free_x"
             , nrow = 1, labeller = label_wrap_gen(width = 15) ) +
  labs(
    title = " PAH concentrations measured in Invertebrates",
    x = "",
    y = "Total PAHs (ng/g)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text( size = 9)
  )
ggsave(filename = "Figures/Fig4b-MeasurementsinInverts.png",    # output file name
  plot = last_plot(),                   # or specify your plot object like `my_plot`
  device = "png",
  width = 8.5,                          # half of standard 11-inch paper width
  height = 4.5,                         # full height of standard paper
  units = "in",
  dpi=300)
#####

## Marine Mammals
#####

D19_mm <- D19 %>%
  filter(Layman_group =="Seals/ Whales") %>%
  mutate(
    logPAH = log(Total_PAH),
    Species_standard = factor(Species_standard)  # Custom facet order
  ) %>%
  group_by(Species_standard, Common_name) %>%
  mutate(Common_name = fct_reorder(Common_name, logPAH, .fun = median, .desc = TRUE)) %>%
  ungroup()


ggplot(D19_mm,  aes(x = Common_name, y = logPAH, color = Common_name)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2, size = 1, alpha = 0.5) +
  facet_wrap(~Species_standard, scales = "free_x"
             , nrow = 1, labeller = label_wrap_gen(width = 15) ) +
  labs(
    title = " Log PAH concentrations measured in Marine Mammals",
    x = "",
    y = "log(Total PAHs (ng/g))"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text( size = 9)
  )
ggsave(filename = "Figures/Fig4c-LogMeasurementsinMarineMammals.png",    # output file name
  plot = last_plot(),                   # or specify your plot object like `my_plot`
  device = "png",
  width = 8.5,                          # half of standard 11-inch paper width
  height = 4.5,                         # full height of standard paper
  units = "in",
  dpi=300)

#NonLog
D19_mm <- D19 %>%
  filter(Layman_group =="Seals/ Whales") %>%
  mutate(
    PAH = Total_PAH,
    Species_standard = factor(Species_standard)  # Custom facet order
  ) %>%
  group_by(Species_standard, Common_name) %>%
  mutate(Common_name = fct_reorder(Common_name, PAH, .fun = median, .desc = TRUE)) %>%
  ungroup()


ggplot(D19_mm,  aes(x = Common_name, y = PAH, color = Common_name)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2, size = 1, alpha = 0.5) +
  facet_wrap(~Species_standard, scales = "free_x"
             , nrow = 1, labeller = label_wrap_gen(width = 15) ) +
  labs(
    title = " PAH concentrations measured in Marine Mammals",
    x = "",
    y = "Total PAHs (ng/g)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text( size = 9)
  )
ggsave(filename = "Figures/Fig4c-MeasurementsinMarineMammals.png",    # output file name
  plot = last_plot(),                   # or specify your plot object like `my_plot`
  device = "png",
  width = 8.5,                          # half of standard 11-inch paper width
  height = 4.5,                         # full height of standard paper
  units = "in",
  dpi=300)

#####

#Figure 5- Time and total PAHs, paneled by taxa - point and line

##Counts of taxa
#####
all_years <- unique(D19$Year_clean)
all_taxa <- unique(D19$Layman_group)
complete_grid <- expand.grid(Year_clean = all_years, Layman_group = all_taxa)

summary_df1 <- D19 %>%
  group_by(Year_clean, Layman_group) %>%
  summarise(Count = n(), .groups = "drop")

summary_df_complete <- complete_grid %>%
  left_join(summary_df1, by = c("Year_clean", "Layman_group")) %>%
  mutate(Count = replace_na(Count, 0)) %>%
  arrange(Year_clean, Layman_group)

summary_cumulative <- summary_df_complete %>%
  group_by(Layman_group) %>%
  arrange(Year_clean) %>%
  mutate(CumulativeCount = cumsum(Count)) %>%
  ungroup()%>%
  mutate(
    Year_clean=as.numeric(Year_clean)
  )

# Plot
ggplot(summary_cumulative, aes(x = Year_clean, y = Count, group = Layman_group, fill= Layman_group)) +
  geom_stream(color = "white", size = 0.1, alpha = 0.9,type = "ridge", bw=1) +
  scale_fill_viridis_d(name = "Taxa",option = "D") +
  theme_minimal(base_size = 13) +
  labs(
    title = "",
    x= "",
    y = "Number of Observations") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = c(0.85, 1.10),  # Position legend inside plot (x, y from bottom-left)
    legend.justification = c("right", "top"),  # Anchor the legend box
    legend.background = element_blank(),   # transparent background
    legend.box.background = element_blank(),  # remove outer box
    legend.key = element_blank(),           # transparent boxes behind keys
    legend.title = element_text(face = "bold"))

ggsave(filename = "Figures/Fig5a-ObservationsByTaxaOverTime.png",    # output file name
  plot = last_plot(),                   # or specify your plot object like `my_plot`
  device = "png",
  width = 8.5,                          # half of standard 11-inch paper width
  height = 5.5,                         # full height of standard paper
  units = "in",
  dpi=300)

#####

## Figure 5- Time trends in concentrations over time
#####
D19_Red <- D19 %>%
  mutate(Year_clean = as.numeric(Year_clean))%>%
  filter(Layman_group%in% c("Fish","Mussels","Clams","Seals/ Whales"))

# Plot
ggplot(D19_Red, aes(x = Year_clean, y = log(Total_PAH))) +
  geom_point(alpha = 0.6, color = "darkblue", size = 1.5) +
  geom_smooth(method = "lm", se = FALSE, color = "firebrick", linewidth = 1) +
  facet_wrap(~ Layman_group, scales = "free_y") +
  labs(
    title = "Log Total PAH Concentrations Over Time by Taxa Group",
    x = "Year",
    y = "Log(Total PAH (ng/g))"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
ggsave(filename = "Figures/Fig5b-LogMeasurementsover Time.png",    # output file name
  plot = last_plot(),                   # or specify your plot object like `my_plot`
  device = "png",
  width = 8.5,                          # half of standard 11-inch paper width
  height = 4.5,                         # full height of standard paper
  units = "in",
  dpi=300)
#####

#Figure 6 - Sample Motivation by Total PAHs by taxa - comparative boxplots
#####
D19a<- D19 %>%
  filter(Layman_group%in% c("Seaweed","Octopus","Crabs/ Shrimp"))

a<-ggplot(data=D19a, aes(x = `Layman_group`, y = log(`Total_PAH`), color = `Sample_motivation`)) +
  geom_point(aes(color = `Sample_motivation` ),position = position_jitterdodge(dodge.width = 0.75,), 
              width = 0.2, height = 0, size = 1, alpha = 0.3) +  # Add jittered points
  geom_boxplot(outlier.shape = NA, fill=NA) +  # Removes the outlier markers from the boxplot
  labs(title = "", x = "", y = "log(Total PAHs (ng/g))") +
  ylim(0,11.5)+
  scale_color_viridis_d() +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none")

D19b<- D19 %>%
  filter(Layman_group%in% c("Snails","Urchins/ Starfish","Fish"))

b<-ggplot(data=D19b, aes(x = `Layman_group`, y = log(`Total_PAH`), color = `Sample_motivation`)) +
  geom_point(aes(color = `Sample_motivation` ),position = position_jitterdodge(dodge.width = 0.75,), 
              width = 0.2, height = 0, size = 1, alpha = 0.3) +  # Add jittered points
  geom_boxplot(outlier.shape = NA, fill=NA) +  # Removes the outlier markers from the boxplot
  labs(title = "", x = "", y = "") +
    ylim(0,11.5)+
  scale_color_manual(values = c("#440154", "#35b779")) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none",
    axis.text.y = element_blank())

D19c<- D19 %>%
  filter(Layman_group%in% c("Clams","Mussels","Seals/ Whales"))

c<-ggplot(data=D19c, aes(x = `Layman_group`, y = log(`Total_PAH`), color = `Sample_motivation`)) +
  geom_point(aes(color = `Sample_motivation` ),position = position_jitterdodge(dodge.width = 0.75,jitter.width = 0.2), 
              width = 0.2, height = 0, size = 1, alpha = 0.3) +  # Add jittered points
  geom_boxplot(outlier.shape = NA, fill=NA) +  # Removes the outlier markers from the boxplot
  labs(title = "", x = "", y = "") +
    ylim(0,11.5)+
  scale_color_viridis_d(name = "Study Motivation") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_blank(),
    legend.position = "right")

ggarrange(a,b,c,widths= c(0.35,0.55,1.7), align = "h",nrow = 1)

ggsave(filename = "Figures/Fig6-LogMeasurementsBySampleMotivation.png",    # output file name
  plot = last_plot(),                   # 
  device = "png",
  width = 10.5,                          # 
  height = 5.5,                         # 
  units = "in",
  dpi=300)

#####

#Figure 7 - Study by Results
#####
study_order <- D19 %>%
  group_by(Study_name_clean) %>%
  summarise(median_PAH = median(log(Total_PAH), na.rm = TRUE)) %>%
  arrange(median_PAH) %>%
  pull(Study_name_clean)

# Step 2: Set Study_name_clean as an ordered factor
D19 <- D19 %>%
  mutate(Study_name_clean = factor(Study_name_clean, levels = study_order),
         Layman_group = factor(Layman_group))

# Step 3: Plot
ggplot(data = D19, aes(x = Study_name_clean, y = log(Total_PAH))) +
  # Jittered points, colored by Taxa group
  geom_jitter(
    aes(color = Layman_group),
    width = 0.2, height = 0,
    size = 1.2, alpha = 0.6
  ) +
  # Boxplot with uniform fill color for each Study (fill used for color)
  geom_boxplot(fill=NA,
    outlier.shape = NA,
    width = 0.6,
    color = "darkgrey",
    alpha = 0.3
  ) +
  # Flip to horizontal
  coord_flip() +

  # Color scales
  scale_color_viridis_d(name = "Taxa Group", option = "D") +

  # Labels
  labs(
    title = "Log PAH Levels in Marine Organisms by Study",
    x = "",
    y = "log(Total PAHs [ng/g])"
  ) +

  # Theme
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8),
    legend.position = "right",
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 9)
  ) +
  ylim(-1, 11.5)

ggsave(filename = "Figures/Fig7-LogMeasurementsByStudy.png",    # output file name
  plot = last_plot(),                   # 
  device = "png",
  width = 8.5,                          # 
  height = 6.5,                         # 
  units = "in",
  dpi=300)
#####
```

Figure 8

Data gaps by regions
scaled to number pie charts for each region for harvest and samples, similar to Wills pie charts on Figure 1 but with size scaling, also explore the idea of having a complimentary table here
#Make a Table on Sampling effort verse Recorded Harvest by Region
```{r}
obs_df <- D20 %>% # Make the Observations data ready to compare to harvest data
  mutate(Taxa= case_when(
    Layman_group=="Seaweed"~"Plants", 
    Layman_group=="Snails"~"Invertebrate", 
    Layman_group=="Urchins/ Starfish"~"Invertebrate", 
    Layman_group=="Octopus"~"Invertebrate", 
    Layman_group=="Clams"~"Invertebrate", 
    Layman_group=="Mussels"~"Invertebrate", 
    Layman_group=="Crabs/ Shrimp"~"Invertebrate", 
    Layman_group=="Fish"~"Fish", 
    Layman_group=="Seals/ Whales"~ "Marine Mammals"
  ))%>%
  group_by(Taxa,Region) %>%
  summarise(
    value = n(),
    .groups = "drop"
  ) %>%
  arrange(Taxa, Region)
subs<- read_csv("subsistence_summary_by_community_coastal.csv") # load harvest data summarized by Will

subs1<-subs%>%
  mutate(Region= case_when(
    is.na(Region) ~ "West",
    TRUE~Region))%>%
  group_by(Region)%>%
  summarise(
    Invertebrate = sum(MarineInverts_kg, na.rm = TRUE),
    Fish = sum(Fish_kg, na.rm = TRUE),
    `Marine Mammals` = sum(MarineMammals_kg, na.rm = TRUE),
    .groups = "drop"
  )

harv_df<-subs1%>%
  pivot_longer(
    cols = c(Invertebrate, Fish, `Marine Mammals`), 
    names_to = "Taxa",           
    values_to = "value"           
 )

library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)
library(stringr)
library(ggforce)


# Add type column to each table
obs_df <- obs_df %>% mutate(type = "Observations")
harv_df <- harv_df %>% mutate(type = "Harvest")

# Rename the value column consistently
colnames(obs_df)[which(!colnames(obs_df) %in% c("Region", "Taxa", "type"))] <- "value"
colnames(harv_df)[which(!colnames(harv_df) %in% c("Region", "Taxa", "type"))] <- "value"

# Combine the two
combined_df <- bind_rows(obs_df, harv_df)

# Check for and remove any rows with missing key data
pie_data <- combined_df %>%
  filter(!is.na(Region), !is.na(Taxa), !is.na(value)) %>%
  group_by(Region, type, Taxa) %>%
  summarise(value = sum(value, na.rm = TRUE), .groups = "drop") %>%
  group_by(Region, type) %>%
  mutate(
    Total = sum(value),
    Fraction = value / Total,
    label = paste0(round(Fraction * 100), "%"),
    ypos = cumsum(Fraction) - 0.5 * Fraction
  ) %>%
  ungroup()

pie_data <- pie_data %>% filter(!is.na(Fraction) & Total > 0)
pie_data<-pie_data%>%#Reorder regionto be more intuitive
  mutate(Region = factor(Region, levels = c("SEAK", "SCAK","Aleutian","West","North")))
## try it with pies scaled to total counts/ kgs

# Prepare pie segments per group
pie_data_summary <- pie_data %>%
  group_by(Region, type, Taxa) %>%
  summarise(value = sum(value, na.rm = TRUE), .groups = "drop") %>%
  group_by(Region, type) %>%
  mutate(
    total = sum(value),
    fraction = value / total,
    start = cumsum(lag(fraction, default = 0)) * 2 * pi,
    end = cumsum(fraction) * 2 * pi,
    middle = (start + end) / 2
  ) %>%
  ungroup()

# Step 2: Normalize radius within each `type`
# Compute the radius relative to max within each `type` (so pies in Harvest are scaled within Harvest group)
radius_scale <- pie_data_summary %>%
  group_by(type, Region) %>%
  summarise(total = unique(total), .groups = "drop") %>%
  group_by(type) %>%
  mutate(radius = sqrt(total) / max(sqrt(total)) * 2)  # scale max radius to 2 units

# Join radius back
pie_data_summary <- pie_data_summary %>%
  left_join(radius_scale, by = c("type", "Region"))

pie_data_summary<-pie_data_summary%>%#Reorder regionto be more intuitive
  mutate(Taxa = factor(Taxa, levels = c("Invertebrate", "Fish","Marine Mammals")))
pie_data_summary<-pie_data_summary%>%#Reorder regionto be more intuitive
  mutate(Region1 = case_when(
    Region=="SEAK"~ "Southeast",
    Region=="SCAK"~ "Southcentral",
    Region=="West"~ "Western",
    Region=="North"~ "Northern",
    TRUE~ Region))
pie_data_summary<-pie_data_summary%>%#Reorder regionto be more intuitive
  mutate(Region1 = factor(Region1, levels = c("Southeast",
"Southcentral", "Aleutian",
"Western",
"Northern")))

write_xlsx(pie_data_summary,"Tables/PieCountsHarvestbyTaxa&Region.xlsx")

# Create the labels and wrap long ones
labels <- c("Observations (scaled by number of observations)", "Reported 2008 Harvest (scaled by kg)")
wrapped_labels <- str_wrap(labels, width = 25)

# Step 3: Plot
ggplot(pie_data_summary) +
  geom_arc_bar(
    aes(
      x0 = as.numeric(factor(Region1)) * 4,
      y0 = ifelse(type == "Observations", 4, 0),
      r0 = 0,
      r = radius,
      start = start,
      end = end,
      fill = Taxa
    ),
    color = "white",
    size = 0.3
  ) +
  coord_fixed() +
  scale_fill_viridis_d(name = "Taxa") +
  theme_void() +
  theme(legend.position = "top") +
  annotate("text",
           x = unique(as.numeric(factor(pie_data_summary$Region1))) * 4,
           y = -2.5, label = unique(pie_data_summary$Region1), size = 3.5,fontface = "bold")+
  annotate("text",
           x = 1.5,
           y = 4.6, 
           label = str_wrap("Observations (scaled by count)", width = 18), 
           size = 3.5,
           angle = 90,  # Rotate 90 degrees
           hjust = 0.5,
           fontface = "bold")+
  annotate("text",
           x = 1.5,
           y = 0, 
           label = str_wrap("Reported 2008 Harvest (scaled by kg)", width = 25), 
           size = 3.5,
           angle = 90,  # Rotate 90 degree
           hjust = 0.5,
    fontface = "bold")

ggsave(filename = "Figures/Fig8-Observations&HarvestComparisonPies.png",    # output file name
  plot = last_plot(),                   # 
  device = "png",
  width = 8.5,                          # 
  height = 6.5,                         # 
  units = "in",
  dpi=300)

```


Notes for Data gaps Tesselation
I would like a full coastal tesselation that highlights where significant knowledge gaps are. I dont necesarily want it to be taxa specific ( looking to "flatten" out all these nuiances) I would like to have a gradient color scale that expands on where high harvest -low harvest and high - low sampling are displayed on the same color scheme, something like a white-blue scale for Harvest and a white to red scale for Sampling all with transparency such that place with relative high for both are purple and those with low for both are white and those with skews are blue or red depending on which direction. Harvest based on total kgs of marine resources harvests regardless of taxa group, sampling effort is number of observations recorded for that area regardless of taxa group

